---
output: beamer_presentation
title: A short How-to for sidora.core
subtitle: Popgen meeting 2022-02-03
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## What is sidora?

- A collection of media and tools to interact with Pandora -- https://github.com/sidora-tools
  - **pandora-the-explorer**: Blog posts exploring Pandora. https://sidora.eva.mpg.de/blog
  - **sidora**: An R shiny app with predefined tables and plots. https://sidora.eva.mpg.de/sidora
  - **pandora2eager**: A script to pump Pandora data into eager2. https://github.com/sidora-tools/pandora2eager
  - ... a number of other things that didn't reach maturity
  - **sidora.core**
  
## What is sidora.core?

- An R package with a number of convenience functions to access Pandora data
- **sidora.core** is also the backbone of **sidora** (webapp) and **pandora2eager** (that's why it's called "core")

## How to install sidora.core on your computer?

- **sidora.core** is not in the CRAN, so it can not be installed with `install.packages()`
- Instead you have to install it from GitHub
  1. Install the package **remotes**, which allows to install R packages from GitHub: `install.packages('remotes')`
  2. Install **sidora.core**: `remotes::install_github("sidora-tools/sidora.core")`
- On Windows you can only install **sidora.core** if you have installed the Rtools (not an R package!) https://cran.r-project.org/bin/windows/Rtools/rtools40.html

## How to connect to the Pandora database server from your computer?

These steps are necessary every time you want to connect to the Pandora server (not the website!):

1. Connect to the institute's subnet via VPN (see the instructions in kbase)
2. Establish an SSH tunnel to the Pandora database server -- and keep it open!

```{bash}
ssh -L 10001:pandora.eva.mpg.de:3306 \
<your username>@daghead1
```

This means that the database server now listens on your computer at `127.0.0.1:10001`.

## How to connect to Pandora with sidora.core?

1. Create a `.credentials` file somewhere on your computer
- This file stores the relevant login information in a way **sidora.core** can understand

```{bash}
<host name of the database server>
<port of the database server>
<database username>
<database password>
```

## How to connect to Pandora with sidora.core?

- The first two entries are what we enabled with the SSH tunnel to the Pandora server
    
```{bash}
127.0.0.1
10001
<database username>
<database password>
```
    
- The last two entries are the credentials of one specific user account, which has the necessary permissions to query data from Pandora
- It's **not your username and password**, but a specific user!
- We can not share this information publicly, but you can get it from Thiseas/James/Stephan/...

## How to connect to Pandora with sidora.core?

2. Create a connection object in R with the `.credentials` file

```{r}
con <- sidora.core::get_pandora_connection(
  "<path to your .credentials file>"
)
```

- This `con` object is what most functions in **sidora.core** need to run and retrieve information from the Pandora database

## How to access Pandora tables with sidora.core?

There are two options:

1. Download the tables as `data.frame`s into your computer's memory and run all computations there

```{r}
sidora.core::get_df("TAB_Site", con) %>%
  dplyr::filter(site.Country == "Germany")
```

- The output of `get_df()` is directly compatible with the tidyverse
- All columns should have the correct data type
- All columns are renamed with an added table prefix: 
    - `Country` -> `site.Country`
    - `Creation_Date` -> `library.Creation_Date`

## How to access Pandora tables with sidora.core?

2. Send queries (dplyr -> SQL) to the database server and only download the results

```{r}
sidora.core::get_con("TAB_Site", con) %>%
  dplyr::filter(Country == "Germany") %>%
  tibble::as_tibble()
```

- IMHO: Option 2 is for advanced users and specific applications
- A lot of the magic of the **sidora.core** package will not work with that: You know what you are doing.

## How to download and merge multiple tables?

Multiple tables can be downloaded with `get_df_list()` and automatically joined with `join_pandora_tables()`

```{r}
sidora.core::get_df_list(c(
  "TAB_Individual", "TAB_Sample"
), con = con) %>%
  sidora.core::join_pandora_tables()
```

## How to download and merge multiple tables?

- This works, because Pandora is strictly hierarchical: Joins always behave the same, and every super-element is present

```{bash}
TAB_Site >
  TAB_Individual >
    TAB_Sample >
      TAB_Extract >
        TAB_Library >
          TAB_Capture >
            TAB_Sequencing >
              TAB_Raw_Data >
                TAB_Analysis
```

- This also allows to automatically bridge the gabs

```{r}
sidora.core::make_complete_table_list(
  c("TAB_Site", "TAB_Capture")
)
```

## How to replace lookup-table keys with real values?

Some values in Pandora are only ids referencing lookup tables (e.g. `Worker`). 

```{r}
sidora.core::get_df("TAB_Library", con) %$% 
  library.Worker %>% unique()
```

*c(22, 4, 15, ...)*

- We can decode them automatically

```{r}
sidora.core::get_df("TAB_Library", con) %>% 
  sidora.core::convert_all_ids_to_values(con) %$% 
  library.Worker %>% unique()
```

*c("Marieke van de Loosdrecht","Alissa Mittnik","Antje Wissgott", ...)*
